---
title: "NRSKAI-phytoplankton"
author: "Catherine Kim"
format: html
editor: source
execute:
  echo: false
  warning: false
---

## Investigating phytoplankton in South Australia

Phytoplankton data from National Reference Station at Kangaroo Island - look for anything interesting in association with suppression of upwelling in SA ~2013.

```{r}
library(planktonr)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(patchwork)
```

## Phytolankton

Get National Referecne Station data using `{planktonr}` package. Look at parameters, column names.

```{r}
pKAI <- pr_get_Indices("NRS", Type = "P") |>
  filter(StationName == "Kangaroo Island") |>
  drop_na() |>
  select(Year_Local, Month_Local, SampleTime_Local, StationName, Parameters, Values) |>
  ungroup()

unique(pKAI$Parameters)
unique(pKAI$StationName)
colnames(pKAI)
```

```{r}
kai <- 

unique(kai$Parameters)
```

Plot all parameters through time.

```{r}
kai |>
  group_split(Parameters) |>
  map(
  ~ ggplot(.x, aes(Year_Local, Values)) +
    geom_point() +
    geom_smooth() +
    scale_x_continuous(n.breaks = 15) +
    labs(title = paste(.x$Parameters), ylab = "Year") +
    theme_classic()
  )
```

Trend analysis

```{r}
pr_plot_Trends(kai |> filter(Parameters == "PhytoBiomassCarbon_pgL"), Trend = "Raw", Survey = "NRS", method = "lm", trans = "log10") /
  pr_plot_Trends(kai |> filter(Parameters == "PhytoAbundance_CellsL"), Trend = "Raw", Survey = "NRS", method = "lm", trans = "log10") /
    pr_plot_Trends(kai |> filter(Parameters == "NoPhytoSpecies_Sample"), Trend = "Raw", Survey = "NRS", method = "lm", trans = "log10") 
```

```{r}
NRSp <- pr_get_Indices("NRS") %>% 
  filter(Parameters == "PhytoAbundance_CellsL") %>% 
  filter(StationCode == "KAI") |>
  drop_na()

unique(NRSp$Parameters)

p1 <- pr_plot_Trends(NRSp, Trend = "Raw", Survey = "NRS", method = "lm", trans = "log10")
p2 <- pr_plot_Trends(NRSp, Trend = "Month", Survey = "NRS", method = "loess")

p1 + p2 + 
  ggplot2::theme(axis.title.y = ggplot2::element_blank()) + # Remove y-title from 2nd column
  patchwork::plot_layout(widths = c(3, 1), guides = "collect")
```

```{r}
(p1 <- planktonr::pr_plot_TimeSeries(NRSp, Survey = "NRS", trans = "log10"))
```
## Functional Group

```{r}
FG <- pr_get_FuncGroups(Survey = "NRS", Type = "P") %>% 
  dplyr::filter(StationCode == "KAI")

phy1 <- planktonr::pr_plot_tsfg(FG, Scale = "Actual")
phy2 <- planktonr::pr_plot_tsfg(FG, Scale = "Actual", Trend = "Month") +
  ggplot2::theme(axis.title.y = element_blank())
    
phy1 + phy2 + 
  patchwork::plot_layout(widths = c(3,1), guides = "collect") & 
  theme(legend.position = "bottom")
```
## Essential Ocean Variables

```{r}
EOV <- planktonr::pr_get_EOVs("NRS")
EOV$Parameters |> unique()

EOV_KAI <- EOV %>%
  filter(StationCode == "KAI") %>%
  pr_remove_outliers(2) %>% 
  pr_get_Coeffs()

EOV_KAI|> filter(Parameters == "Oxygen_umolL")
```

```{r}
pr_plot_EOVs(EOV_KAI, EOV = "Biomass_mgm3") /
  pr_plot_EOVs(EOV_KAI, EOV = "PhytoBiomassCarbon_pgL", col = "darkgreen") /
  pr_plot_EOVs(EOV_KAI, EOV = "CTDTemperature_degC", col = "darkblue") /
  pr_plot_EOVs(EOV_KAI, EOV = "ShannonCopepodDiversity", col = "bisque") &
    theme_bw() &
  plot_annotation(title = "Kangaroo Island")

pr_plot_EOVs(EOV_KAI, EOV = "ShannonPhytoDiversity", col = "burlywood") /
  pr_plot_EOVs(EOV_KAI, EOV = "Salinity", col = "turquoise") /
  pr_plot_EOVs(EOV_KAI, EOV = "PigmentChla_mgm3", col = "slateblue4") /
  pr_plot_EOVs(EOV_KAI, EOV = "Ammonium_umolL", col = "tomato4") &
    theme_bw() 

pr_plot_EOVs(EOV_KAI, EOV = "Nitrate_umolL", col = "violetred3") /
  pr_plot_EOVs(EOV_KAI, EOV = "Nitrite_umolL", col = "moccasin") /
  pr_plot_EOVs(EOV_KAI, EOV = "Silicate_umolL") / 
  pr_plot_EOVs(EOV_KAI, EOV = "Phosphate_umolL", col = "olivedrab") &
#  pr_plot_EOVs(EOV_KAI, EOV = "Oxygen_umolL", col = "darkred") & 
  theme_bw()
```

```{r eval=FALSE}
EOV_KAI |>
  group_split(Parameters) |>
  map(~ pr_plot_EOVs(.x, EOV = .x$Parameters))
```

## Zooplankton

```{r}
NRSz <- planktonr::pr_get_Indices("NRS", "Z") %>% 
  filter(Parameters == "ZoopAbundance_m3") %>% 
  filter(StationCode %in% c("KAI")) |>
  drop_na()

p1 <- planktonr::pr_plot_Trends(NRSz, Trend = "Raw", Survey = "NRS", method = "lm", trans = "log10")
p2 <- planktonr::pr_plot_Trends(NRSz, Trend = "Month", Survey = "NRS", method = "loess")

p1 + p2 + 
  ggplot2::theme(axis.title.y = ggplot2::element_blank()) + # Remove y-title from 2nd column
  plot_layout(widths = c(3, 1), guides = "collect")
```

Climatologies

```{r}
(p1 <- planktonr::pr_plot_TimeSeries(NRSz, Survey = "NRS", trans = "log10"))
(p2 <- planktonr::pr_plot_Climatology(NRSz, Survey = "NRS", Trend = "Month", trans = "log10"))
(p3 <- planktonr::pr_plot_Climatology(NRSz, Survey = "NRS", Trend = "Year", trans = "log10"))

p1 / p2/ p3
```

Functional Groups

```{r}
FGz <- pr_get_FuncGroups(Survey = "NRS", Type = "Z") %>% 
  filter(StationCode == "KAI")

z1 <- planktonr::pr_plot_tsfg(FGz, Scale = "Actual")
z2 <- planktonr::pr_plot_tsfg(FGz, Scale = "Actual", Trend = "Month") +
  ggplot2::theme(axis.title.y = element_blank())

(phy1 + phy2) / 
(z1 + z2) + 
  patchwork::plot_layout(widths = c(3,1), guides = "collect") & 
  theme(legend.position = "bottom")
```

